Measuring contrast processing in the visual system using SSVEP

### 2024-05-08 review.qmd
Two more

Later in the visual pathway, we would like to know more about the role of corticothalamic feedback in the LGN. Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@Briggs2007; @Sherman1996; @OConnor2002; @Gouws2014]. Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular, but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@Baker2007; @Dougherty2019] to begin even at this relatively early stage. Frequency tagging inputs to different eyes or in different precortical pathways allows us to address neurons in different parts of the LGN. These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012; @Attal2013; @Tesche1996], noninvasive deep-brain stimulation techniques [@MohammadjavadietalSciRep2022] or implanted electrode arrays [@Krolak-Salmon2003]) may allow us to study these computations in more detail.

Finally, SSVEPs continue to provide a way of studying contrast in the cortex. Here, we are often interested in how different visual parameters interact. For example, how are signals from different eyes combined to generate both scalar contrast values and also binocular depth cues? How does contrast combination depend on the low level properties of the individual inputs such as their retinotopic location, cone contrast, spatiotemporal frequency, eye of origin and orientation? Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@Li22; @Stoughton2008; @Kaneko2020; @Gegenfurtner2003; @Solomon2007; @Wandell1993]. All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages [@Watts2024; @Baker2017; @Chen2021; @Katyal2018; @Busse2009]. The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each lcoation provided detailed information about the processing nonlinearities up to and including that point. In principle, this would allow researchers to uniquely fit the parameters of their computations models of visual processing [@Regan1988].  To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the inputs frequencies or simple sums and differences of those frequencies). As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@Chariker2020; @Groen2022; @Baker2017; @Tsai2012; @Schrimpf2020]. Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subject. In principle, we are therefore able to use the SSVEP to fit the parameters of early visual processing. This approach may also allow us to develop more sensitive tests for the changes in early visual processing that accompany a wide range of neurological diseases and disorders.

>>>>

Later in the visual pathway, we would like to know more about the role of corticothalamic feedback in the LGN. Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@Briggs2007; @Sherman1996; @OConnor2002; @Gouws2014]. Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular, but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@Baker2007; @Dougherty2019] to begin even at this relatively early stage. Frequency tagging inputs to different eyes or in different precortical pathways allows us to address neurons in different parts of the LGN. These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012; @Attal2013; @Tesche1996], noninvasive deep-brain stimulation techniques [@Mohammadjavadi2022] or implanted electrode arrays [@Krolak-Salmon2003]) may allow us to study these computations in more detail.

Finally, SSVEPs continue to provide a way of studying contrast in the cortex. Here, we are often interested in how different visual parameters interact. For example, how are signals from different eyes combined to generate both scalar contrast values and also binocular depth cues? How does contrast combination depend on the low level properties of the individual inputs such as their retinotopic location, cone contrast, spatiotemporal frequency, eye of origin and orientation? Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@Li2022; @Stoughton2008; @Kaneko2020; @Gegenfurtner2003; @Solomon2007; @Wandell1993]. All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages [@Watts2024; @Baker2017; @Chen2021; @Katyal2018; @Busse2009]. The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each lcoation provided detailed information about the processing nonlinearities up to and including that point. In principle, this would allow researchers to uniquely fit the parameters of their computations models of visual processing [@Regan1988].  To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the inputs frequencies or simple sums and differences of those frequencies). As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@Chariker2020; @Groen2022; @Baker2017; @Tsai2012; @Schrimpf2020]. Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subject. In principle, we are therefore able to use the SSVEP to fit the parameters of early visual processing. This approach may also allow us to develop more sensitive tests for the changes in early visual processing that accompany a wide range of neurological diseases and disorders.
